---
title: "Ocean acidification affects microbial community and invertebrate settlement on biofilms:Fig2"
author: "Sergio E. Morales"
date: "26 June 2017"
---

###Pre-processing summary
  All data was processed in Qiime, clustered at 97% sequence similarity and assigned taxonomy against SILVA 119. Libraries were subsampled and rarefied (to 8000 reads) ten times.  OTU tables generated were ten combined and the new data used for downstream analysis in R as shown below.


```{r Load libraries and relative abundance data for invertebrates}

library(ggplot2)
library(vegan)
library(MASS)
library(ade4)
library(grid)

taxa <- read.csv("/file path/Github/Figure 2/Tax_rel_ab.csv",fill = TRUE, header = TRUE, sep = ",")
dim(taxa)
row.names(taxa) <- taxa$SampleID
taxa <- taxa[, -1]
data.prop <- taxa/rowSums(taxa)
data.prop[1:3, 1:3]

# Calculate Shannon index and save results to table
shann <- diversity(taxa)  
shann
write.table(shann, file = "/file path/Github/Figure 2/shann.txt", sep = "\t") #added shannon results to data frame rename MD below.

MD <- read.csv("/file path/Github/Figure 2/Fig2_taxa_MD.csv")

#Calculate ANOVA/Tukeys HSD for invertebrate Shannon diversity by pH
aov.Shan_inv.pH = aov(Shann ~ pH_Treatment, data = MD)
summary(aov.Shan_inv.pH)
TukeyHSD(aov.Shan_inv.pH)
```

###NMDS ordination using invertebrate taxa abundance
```{r Make NMDS plot macro taxa, results='markup'}

inv_MDS <- metaMDS(data.prop)
inv_MDS
inv_MDS <- metaMDS(data.prop, previous = inv_MDS)
inv_MDS
plot(inv_MDS)

NMDS = data.frame(MDS1 = inv_MDS$points[,1], MDS2 = inv_MDS$points[,2],pH_Treatment=MD$pH_Treatment)

ord<-ordiellipse(inv_MDS, NMDS$pH_Treatment, display = "sites", 
                 kind = "se", conf = 0.95, label = T)

veganCovEllipse <-
  function(cov, center = c(0,0), scale = 1, npoints = 100)
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    Q <- chol(cov, pivot = TRUE)
    o <- attr(Q, "pivot")
    t(center + scale * t(Circle %*% Q[,o]))
  }

df_ell <- data.frame()
for(g in levels(NMDS$pH_Treatment)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$pH_Treatment==g,],
                                                   veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,pH_Treatment=g))
}


my_grob = grobTree(textGrob("Stress = 0.02", x=0.6,  y=0.95, hjust=0,
                            gp=gpar(col="Black", fontsize=14, fontface="italic"))) 


#Plot Fig 2a
Fig2a_macro<-ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(size=5,aes(color = pH_Treatment)) +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=pH_Treatment), size=1, linetype=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x  = element_text( vjust=0.5, size=12)
  )+
  theme(axis.title.x = element_text(face="bold",size=14),
        axis.text.x = element_text(colour = "black", vjust=1, hjust = 1, size=14),
        axis.text.y = element_text(colour = "black", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=14),
        legend.text = element_text(size = 14),
        legend.position="right",
        #Manipulating the facet features
        strip.text.x = element_text(size=12, face="bold"),
        strip.text.y = element_text(size=12, face="bold"),
        strip.background = element_rect(colour="black")) + 
  guides(fill = guide_legend(ncol = 1, title.hjust = 0.4)) + annotation_custom(my_grob) + scale_colour_grey() + guides(colour=guide_legend(title="pH"))
Fig2a_macro

ggsave("/file path/Github/Figure 2/Fig2A.pdf",units = c("in"), width=6,height=4)

#Plot stressplot
stressplot(inv_MDS)
dev.copy(pdf, "/file path/Github/Figure 2/FigS1.pdf", width=8,height=4)
dev.off()

```

#Calculate ANOSIM for invertebrate community against pH treatment
```{r ANOSIM macro}
macro.dist <- vegdist(data.prop)
attach(MD)
macro.ano <- anosim(macro.dist, pH_Treatment)
summary(macro.ano)
plot(macro.ano)
dev.copy(pdf, "/file path/Github/Figure 2/ANOSIM_macro.pdf", width=6,height=4)
dev.off()
```

```{r}
(.packages())
detachAllPackages <- function() {
  
  basic.packages <- c("package:stats","package:graphics","package:grDevices","package:utils","package:datasets","package:methods","package:base")
  
  package.list <- search()[ifelse(unlist(gregexpr("package:",search()))==1,TRUE,FALSE)]
  
  package.list <- setdiff(package.list,basic.packages)
  
  if (length(package.list)>0)  for (package in package.list) detach(package, character.only=TRUE)
  
}

detachAllPackages()
```

```{r Load libraries and amplicon data}
library(phyloseq)  
library(ggplot2)
library(plyr)
library(scales)
library(reshape)
library(RColorBrewer)
library(grid)
library(vegan)
library(ggrepel)

otutable_biom_file <- paste("/file path/Github/", "merged_otu_table.biom", sep = "")
map_file <- paste("/file path/Github/Figure 2/", "Nelson_etal_Fig2_map.txt", sep = "")
```

```{r Create phyloseq object}
biom_otu_tax <- import_biom(otutable_biom_file)
bmsd <- import_qiime_sample_data(map_file)
biom_otu_tax

KN_phyloseq <- merge_phyloseq(biom_otu_tax, bmsd)
KN_phyloseq
sample_sums(KN_phyloseq)
```

#####Create average result for multiple rarefaction by transforming data using (divide by 10) and check counts per sample
```{r Create average result for multiple rarefaction by transforming data using (divide by 10), results='markup'}
KN_phyloseq = transform_sample_counts(KN_phyloseq, function(x) x/10)
sample_sums(KN_phyloseq)
```

##### Round and confirm count number
```{r Round and confirm count number, results='markup'}
KN_phyloseq = transform_sample_counts(KN_phyloseq, round)
sample_sums(KN_phyloseq)
KN_phyloseq = prune_samples(sample_sums(KN_phyloseq)>=1, KN_phyloseq)
sample_sums(KN_phyloseq)
```

#####Prune taxa with less than 1 count
```{r  Save original file and create new file with only present (no zeroes) taxa, results='markup', echo=TRUE}

#Create new file with only present (no zeroes) taxa
KN_phyloseq = prune_taxa(taxa_sums(KN_phyloseq) > 1, KN_phyloseq)
any(sample_sums(KN_phyloseq) == 0)
any(sample_sums(KN_phyloseq) > 0)
sum(taxa_sums(KN_phyloseq) > 0)
any(sample_sums(KN_phyloseq) < 1)
sum(taxa_sums(KN_phyloseq) < 1)
```

###Compare sequences per sample or OTU
```{r Compare sequences per sample or OTU}
readsumsdf = data.frame(nreads = sort(taxa_sums(KN_phyloseq), TRUE), sorted = 1:ntaxa(KN_phyloseq), 
                        type = "OTU")
readsumsdf = rbind(readsumsdf, data.frame(nreads = sort(sample_sums(KN_phyloseq), 
                                                        TRUE), sorted = 1:nsamples(KN_phyloseq), type = "Samples"))
title = "Total number of reads"
p = ggplot(readsumsdf, aes(x = sorted, y = nreads)) + geom_bar(stat = "identity")
p + ggtitle(title) + scale_y_log10() + facet_wrap(~type, 1, scales = "free")
```

```{r}
sample_variables(KN_phyloseq)
```

```{r Attached OTU ID}
tax_table(KN_phyloseq) <- cbind(tax_table(KN_phyloseq), OTU=taxa_names(KN_phyloseq))
```

```{r Rename Ranks}
colnames(tax_table(KN_phyloseq)) = c("Domain", "Phylum", "Class", "Order", "Family", "Genus", "Species", "OTU")
```

```{r Calculate Alpha diversity and anova for all micro data}
alpha_summary_all <- estimate_richness(KN_phyloseq, measures = c("Shannon", "Observed"))
#combine with metadata
alpha_meta_all <- data.frame(alpha_summary_all, sample_data(KN_phyloseq))
alpha_meta_all$Biofilm_Age_days<-as.factor(alpha_meta_all$Biofilm_Age_days)

#Calculate ANOVA/Tukeys HSD for Shannon by pH and Age 
aov.Shan_micro.pH_age = aov(Shannon ~ pH*Biofilm_Age_days, data = alpha_meta_all)
summary(aov.Shan_micro.pH_age)
TukeyHSD(aov.Shan_micro.pH_age)

#Calculate ANOVA/Tukeys HSD for Richness by pH and Age 
aov.Rich_micro.pH_age = aov(Observed ~ pH*Biofilm_Age_days, data = alpha_meta_all)
summary(aov.Rich_micro.pH_age)
TukeyHSD(aov.Rich_micro.pH_age)
```


### Plot NMDS ordination Bray Distance for all micro data
```{r NMDS ordination, results='markup'}
NMDSord <- ordinate(KN_phyloseq, "NMDS", "bray")
NMDSord


sample_data(KN_phyloseq)$Biofilm_Age_days = factor( get_variable(KN_phyloseq, "Biofilm_Age_days"))
       
plot_ordination(KN_phyloseq, NMDSord, type = "samples", color = "pH_Treatment", shape = "Biofilm_Age_days") + geom_point(size = 5) + geom_jitter()

sampleplot = plot_ordination(KN_phyloseq, NMDSord, type = "samples", color = "pH_Treatment", shape = "Biofilm_Age_days") 

my_grob = grobTree(textGrob("Stress = 0.11", x=0.6,  y=0.95, hjust=0,
                            gp=gpar(col="Black", fontsize=14, fontface="italic")))




NMDS_all = sampleplot + geom_point(size = 5) + theme(text = element_text(size = 16)) + theme(axis.title.x = element_text(face="bold",size=16),
                                          axis.text.x = element_text(angle=0, colour = "black", vjust=1, hjust = 1, size=14),
                                          axis.text.y = element_text(colour = "black", size=14),
                                          axis.title.y = element_text(face="bold",size=14),
                                          plot.title = element_text(size = 18),
                                          legend.title =element_text(size = 14),
                                          legend.text = element_text(size = 14),
                                          legend.position="right",
                                          legend.key.size = unit(0.50, "cm"),
                                          strip.text.x = element_text(size=12, face="bold"),
                                          strip.text.y = element_text(size=12, face="bold"),
                                          panel.background = element_blank(),
                                          panel.border = element_rect(fill = NA, colour = "black"),
                                          strip.background = element_rect(colour="black")) + annotation_custom(my_grob)

NMDS_all_plot= NMDS_all + geom_jitter() 
NMDS_all_plot
ggsave("/file path/Github/Figure 2/FigS2.pdf",units = c("in"), width=6,height=4)

stressplot(NMDSord)
dev.copy(pdf, "/file path/Github/Figure 2/FigS4.pdf", width=8,height=4)
dev.off()

```

#####Subset samples by Age and only keep 27 day old data
```{r Subset samples by young biofilm only}
Age27_phyloseq = subset_samples(KN_phyloseq, Biofilm_Age_days == "27")
```

### Plot NMDS ordination Bray Distance (only 27days)
```{r NMDS ordination for young biofilm, results='markup'}
NMDSord27 <- ordinate(Age27_phyloseq, "NMDS", "bray")
NMDSord27


plot_ordination(Age27_phyloseq, NMDSord27, type = "samples", color = "pH_Treatment") + geom_point(size = 5) + geom_jitter()
baseplot = plot_ordination(Age27_phyloseq, NMDSord27, type = "samples", color = "pH_Treatment") 

my_grob = grobTree(textGrob("Stress = 0.05", x=0.6,  y=0.95, hjust=0,
                            gp=gpar(col="Black", fontsize=14, fontface="italic")))




NMDSord27_plot = baseplot + geom_point(size = 5) + theme(text = element_text(size = 16)) + theme(axis.title.x = element_text(face="bold",size=16),
                                                                                      axis.text.x = element_text(angle=0, colour = "black", vjust=1, hjust = 1, size=14),
                                                                                      axis.text.y = element_text(colour = "black", size=14),
                                                                                      axis.title.y = element_text(face="bold",size=14),
                                                                                      plot.title = element_text(size = 18),
                                                                                      legend.title =element_text(size = 14),
                                                                                      legend.text = element_text(size = 14),
                                                                                      legend.position="right",
                                                                                      legend.key.size = unit(0.50, "cm"),
                                                                                      strip.text.x = element_text(size=12, face="bold"),
                                                                                      strip.text.y = element_text(size=12, face="bold"),
                                                                                      panel.background = element_blank(),
                                                                                      panel.border = element_rect(fill = NA, colour = "black"),
                                                                                      strip.background = element_rect(colour="black")) + annotation_custom(my_grob)

Fig2b= NMDSord27_plot + geom_jitter()# pass this if you want labels + geom_text_repel(aes(label = SampleID), size = 5) 
Fig2b_v1
ggsave("/file path/Github/Figure 2/Fig2B_only27days.pdf",units = c("in"), width=6,height=4)

stressplot(NMDSord27)
dev.copy(pdf, "/file path/Github/Figure 2/FigS5.pdf", width=8,height=4)
dev.off()

```


### Plot NMDS ordination Bray Distance with ellipses (only 27days)
```{r NMDS ordination with ellipses only young biofilms, results='markup'}

points <- as.data.frame(NMDSord27$points)

NMDS = data.frame(MDS1 = points$MDS1, MDS2 = points$MDS2)
NMDS$pH <- get_variable(Age27_phyloseq, "pH")
plot(NMDSord27$points, col = NMDS$pH)

ord<-ordiellipse(NMDSord27, NMDS$pH, display = "sites", 
                 kind = "se", conf = 0.95, label = T)

veganCovEllipse <-
  function(cov, center = c(0,0), scale = 1, npoints = 100)
  {
    theta <- (0:npoints) * 2 * pi/npoints
    Circle <- cbind(cos(theta), sin(theta))
    Q <- chol(cov, pivot = TRUE)
    o <- attr(Q, "pivot")
    t(center + scale * t(Circle %*% Q[,o]))
  }

df_ell <- data.frame()
for(g in levels(NMDS$pH)){
  df_ell <- rbind(df_ell, cbind(as.data.frame(with(NMDS[NMDS$pH==g,],
                                                   veganCovEllipse(ord[[g]]$cov,ord[[g]]$center,ord[[g]]$scale)))
                                ,pH=g))
}

my_grob = grobTree(textGrob("Stress = 0.05", x=0.6,  y=0.95, hjust=0,
                            gp=gpar(col="Black", fontsize=14, fontface="italic")))


Fig2b_ellipses<-ggplot(data = NMDS, aes(MDS1, MDS2)) + geom_point(size=5,aes(color = pH)) +
  geom_path(data=df_ell, aes(x=NMDS1, y=NMDS2,colour=pH), size=1, linetype=2)+
  theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x  = element_text( vjust=0.5, size=12)
  )+
  theme(axis.title.x = element_text(face="bold",size=14),
        axis.text.x = element_text(colour = "black", vjust=1, hjust = 1, size=14),
        axis.text.y = element_text(colour = "black", size=14),
        axis.title.y = element_text(face="bold", size=14),
        legend.title = element_text(size=14),
        legend.text = element_text(size = 14),
        legend.position="right",
        #Manipulating the facet features
        strip.text.x = element_text(size=12, face="bold"),
        strip.text.y = element_text(size=12, face="bold"),
        strip.background = element_rect(colour="black")) + # Black rectangle around facet title
  guides(fill = guide_legend(ncol = 1, title.hjust = 0.4)) + annotation_custom(my_grob) + scale_colour_grey() + guides(colour=guide_legend(title="pH"))
Fig2b_ellipses

ggsave("/file path/Github/Figure 2/Fig2B.pdf",units = c("in"), width=6,height=4)

stressplot(NMDSord27)
dev.copy(pdf, "/file path/Github/Figure 2/Fig2B_ellipse_stress.pdf", width=8,height=4)
dev.off()

```

```{r ANOSIM or ADONIS micro data}
#all data
df = as(sample_data(KN_phyloseq), "data.frame")
d = distance(KN_phyloseq, "bray")
pH_Age_all_adonis = adonis2(d ~ pH_all*Age_all, df)
pH_Age_all_adonis


#27days
pH_27days = get_variable(Age27_phyloseq, "pH")
pH_27days_ano = anosim(distance(Age27_phyloseq, "bray"), pH_27days)
summary(pH_27days_ano)
pH_27days_ano$signif
pH_27days_ano$statistic
plot(pH_27days_ano)
dev.copy(pdf, "/file path/Github/Figure 2/ANOSIM_micro.pdf", width=6,height=4)
dev.off()
```


```{r Multi plot function}
multiplot <- function(..., plotlist = NULL, file, cols = 1, layout = NULL) {
  require(grid)

  plots <- c(list(...), plotlist)

  numPlots = length(plots)

  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                 ncol = cols, nrow = ceiling(numPlots/cols))
}

if (numPlots == 1) {
print(plots[[1]])

} else {
grid.newpage()
pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))

for (i in 1:numPlots) {
  matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))

  print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                  layout.pos.col = matchidx$col))
 }
}
 }
```

```{r Multiplot NMDS}
pdf("/file path/Github/Figure 2/Fig2_multinmds.pdf",width=6,height=8)
multiplot(Fig2a_macro,Fig2b_ellipses, cols=1)
dev.off()
```

```{r extract taxonomy info as matrix for 27 days only}
taxonomy = tax_table(Age27_phyloseq, errorIfNULL=FALSE)
if( !is.null(taxonomy) ){
  taxonomy = data.frame(as(taxonomy, "matrix"))
} 

otus <- otu_table(Age27_phyloseq)
otus_mat <- otus@.Data
```

###Determine number of clusters for KN using elbow graph 27 days only
```{r Determine number of clusters for KN using elbow graph 27 days only}
# Determine number of clusters
wss <- (nrow(otus_mat)-1)*sum(apply(otus_mat,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(otus_mat, 
                                     centers=i)$withinss)
plot(1:15, wss, type="b", xlab="Number of Clusters",
     ylab="Within groups sum of squares")
```
#####Results suggest 2-4 clusters
Ran simprof for 27 days only choosing expected cluster number by NMDS results and elbow grpah, plot as color coded tree with Bray Curtis Distance.  

###Test clustering for KN using SIMPROF 27 days only
```{r SIMPROF tree for KN with samples clustered by color 27 days only}
library(clustsig)
library(vegan)
spistSIMPROF <- simprof(otus_mat, num.expected=4, num.simulated=3, method.cluster='average', method.distance='braycurtis', method.transform='identity', alpha=0.05, sample.orientation='column', silent=FALSE)

simprof.plot(spistSIMPROF, leafcolors=NA, plot=TRUE, fill=TRUE, leaflab="perpendicular", siglinetype=1)
```
#####Results suggest ~2 main branches
Used a second clustering method which allows for both bootstrapping, and confidence determination.  


###Pvclust tree for KN using bray distance and including p values for each node. [AU (approximately unbiased) BP (bootstrap probability)]. Boxes mark clusters with 90% confidence 27 days only
```{r Cluster tree for KN with box marking 90percent confidence 27 days only}
library(pvclust)
library(magrittr)
library(dendextend)
library("devtools")
source_url("https://raw.githubusercontent.com/nielshanson/mp_tutorial/master/taxonomic_analysis/code/pvclust_bcdist.R")
fit <- pvclust(otus_mat, method.hclust="ward", method.dist="bray–curtis", n=1000)

plot(fit) 

dend <- as.dendrogram(fit)
fit %>% as.dendrogram %>% 
  plot(main = "Cluster dendrogram with AU/BP values (%)\n reproduced plot with dendrogram")
fit %>% text
fit %>% pvrect
pvrect(fit, alpha=.90)

labels(dend)

require(colorspace)

pH<-as.factor(bmsd$pH_Treatment)
pH_Treatment<-data.frame(bmsd$pH_Treatment)
colnames(pH_Treatment) <- "pH_Treatment"
col<-pH


colors_to_use <- as.numeric(pH)
colors_to_use
colors_to_use <- colors_to_use[order.dendrogram(dend)]
colors_to_use

labels_colors(dend) <- colors_to_use
labels_colors(dend)
labels_colors(dend) <- rainbow_hcl(3)[as.numeric(col)]

labels_colors(dend) 
plot(dend) 
legend("topleft", legend = levels(pH), fill = rainbow_hcl(3), cex=0.75)
pvrect(fit, alpha=.90)
dev.copy(pdf, "/file path/Github/Figure 2/FigS6.pdf", width=12,height=6)
dev.off()
```
#####Results at 90% suggest 2 clusters.


```{r Calculate Alpha diversity and anova for micro using young data}
alpha_summary <- estimate_richness(Age27_phyloseq, measures = c("Shannon", "Observed"))
#combine with metadata
alpha_meta <- data.frame(alpha_summary, sample_data(Age27_phyloseq))

#Calculate ANOVA/Tukeys HSD for Shannon by pH
aov.Shan_micro.pH = aov(Shannon ~ pH, data = alpha_meta)
summary(aov.Shan_micro.pH)
TukeyHSD(aov.Shan_micro.pH)

#Calculate ANOVA/Tukeys HSD for Richness by pH
aov.Rich_micro.pH = aov(Observed ~ pH, data = alpha_meta)
summary(aov.Rich_micro.pH)
TukeyHSD(aov.Rich_micro.pH)
```


```{r Plot Shannon for invertebrates}
macro_alpha<-ggplot(MD, aes(x = pH_Treatment, y=Shann, color=pH_Treatment)) + theme(text = element_text(size = 14)) + geom_boxplot()

Fig2c<-macro_alpha + theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x  = element_text( vjust=0.5, size=12)
  ) + theme(axis.title.x = element_text(face="bold",size=14),
        axis.text.x = element_text(colour = "black", vjust=1, hjust = 1, size=14, angle = 0),
        axis.text.y = element_text(colour = "black", size=14),
        axis.title.y = element_text(face="bold", size=14),
        #Manipulating the facet features
        strip.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text.y = element_blank(),
        legend.position="none") + scale_colour_grey() +
    xlab("pH") +    
    ylab("Shannon") 
Fig2c

ggsave("/file path/Github/Figure 2/Fig2c.pdf",width=4,height=4)
```



```{r Plot Shannon micro data only 27 days}
micro_alpha= plot_richness(Age27_phyloseq, x = "pH", color="pH", measures = c("Shannon")) +theme(text = element_text(size = 14)) + geom_boxplot() 
  
  
Fig2d<-micro_alpha + theme_bw() +
  theme(
    plot.background = element_blank()
    ,panel.grid.major = element_blank()
    ,panel.grid.minor = element_blank()
    ,panel.background = element_blank()
    ,axis.text.x  = element_text( vjust=0.5, size=12)
  ) + theme(axis.title.x = element_text(face="bold",size=14),
        axis.text.x = element_text(colour = "black", vjust=1, hjust = 1, size=14, angle = 0),
        axis.text.y = element_text(colour = "black", size=14),
        axis.title.y = element_text(face="bold", size=14),
        #Manipulating the facet features
        strip.text.x = element_blank(),
        strip.background = element_blank(),
        strip.text.y = element_blank(),
        legend.position="none") + scale_colour_grey() +
    xlab("pH") +    
    ylab("Shannon") 
Fig2d

ggsave("/file path/Github/Figure 2/Fig2d.pdf",width=4,height=4)
```

```{r Multiplot alpha}
pdf("/file path/Github/Figure 2/alpha.pdf",width=4,height=8)
multiplot(Shan_macro, Alpha27, cols=1)
dev.off()

```

###Rel abundance for macroorganisms
```{r Stacked bar rel abund macro}
library(MASS)
library(ade4)
library(reshape)
library(Rmisc)

taxa <- read.csv("/file path/Github/Figure 2/Macro_abund.csv",fill = TRUE, header = TRUE, sep = ",")
dim(taxa)
row.names(taxa) <- taxa$SampleID
taxa <- taxa[, -1]
taxa_rel_ab <- taxa/rowSums(taxa)
taxa_rel_ab[1:3, 1:3]


var <- read.csv("/file path/Github/Figure 2/var.csv",fill = TRUE, header = TRUE, sep = ",")
row.names(var) <- var$SampleID

tax_ra_df <- merge(taxa_rel_ab,var,by="row.names")

tax_ra_df <- melt(tax_ra_df, id=c("SampleID","pH_Treatment"))
names(tax_ra_df)[names(tax_ra_df)=="variable"] <- "Taxa"
tax_ra_df = tax_ra_df[-(1:12),]
as.numeric(tax_ra_df$value)
names(tax_ra_df)[names(tax_ra_df)=="value"] <- "Relative Abundance"
as.factor(tax_ra_df$pH_Treatment)
as.factor(tax_ra_df$Taxa)
write.csv(tax_ra_df, file="/file path/Github/Figure 2/tax_ra_df.csv", row.names=TRUE)

tax_ra_df <- read.csv("/file path/Github/Figure 2/tax_ra_df.csv",fill = TRUE, header = TRUE, sep = ",")


SSE_macro <- summarySE(tax_ra_df, measurevar="Relative.Abundance", groupvars=c("pH_Treatment","Taxa"))
SSE_macro
write.csv(tax_ra_df, file="/file path/Github/Figure 2/tax_SSE_macro_df.csv", row.names=TRUE, col.names=NA)


d= ggplot(SSE_macro,aes(x=pH_Treatment,y=Relative.Abundance,fill=Taxa))+ geom_bar(stat="identity",position="fill") +  scale_y_continuous(labels = percent_format())

c= d+ xlab("pH")

b= c+ ylab("Relative Abundance")

a= b+ scale_fill_manual(name="Taxa",values = c("grey26", "royalblue", "chartreuse3",  "red", "darkorange","cyan2", "darkgreen", "deepskyblue", "mediumorchid3","#89C5DA", "#DA5724", "#74D944", "#C84248", "#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030",   "#AD6F3B", "#CD9BCD", "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#8569D5", "#5E738F", "#D1A33D", "#8A7C64", "#599861", "blue4", "yellow1", "violetred", "#990000", "#99CC00", "#003300", "#00CCCC", "#9966CC", "#993366", "#990033", "#4863A0", "#000033", "#330000", "#00CC99", "#00FF33", "#00CCFF", "#FF9933", "#660066", "#FF0066", "#330000", "#CCCCFF", "#3399FF", "#66FFFF", "#B5EAAA","#FFE87C"))

z= a+ theme(axis.title.x = element_text(face="bold",size=14),
            axis.text.x = element_text(angle=0, colour = "black", vjust=1, hjust = 1, size=14),
            axis.text.y = element_text(colour = "black", size=14),
            axis.title.y = element_text(face="bold",size=14),
            plot.title = element_text(size = 14),
            legend.title =element_text(face="bold",size = 14),
            legend.text = element_text(size = 14),
            legend.position="right",
            legend.key.size = unit(0.50, "cm"),
            strip.text.x = element_text(size=12, face="bold"),
            strip.text.y = element_text(size=12, face="bold"),
            panel.background = element_blank(),
            panel.border = element_rect(fill = NA, colour = "black"),
            strip.background = element_rect(colour="black"))

Fig2e= z+ guides(fill = guide_legend(reverse = TRUE, ncol=1))
Fig2e

ggsave("/file path/Github/Figure 2/Fig2e.pdf",width=6,height=4)
```


### Plot micro phylum by pH_Treatment percent transformed 27 days only
###OTUs at Phylum level relative abundance with only abundant groups
```{r OTUs at Phylum level rel abund with no OTU breaks only abundant groups}
require("genefilter")

Age27_phylum_glom <- tax_glom(Age27_phyloseq, taxrank = "Phylum")
tax_table(Age27_phylum_glom) =gsub("D_1__", "", tax_table(Age27_phylum_glom))

Age27_phylum_glom  = transform_sample_counts(Age27_phylum_glom, function(x) x / sum(x) )
Age27_phylum_glom_df  = psmelt(Age27_phylum_glom)
flist    <- filterfun(kOverA(8, 0.0001))
PNG.logi <- filter_taxa(Age27_phylum_glom, flist)
PNG.trim <- filter_taxa(Age27_phylum_glom, flist, TRUE)
Age27_phylum_glom_MA <- identical(PNG.trim, prune_taxa(PNG.logi, Age27_phylum_glom)) 
Age27_phylum_glom_MA <- identical(sum(PNG.logi), ntaxa(PNG.trim))
Age27_phylum_glom_MA <- filter_taxa(Age27_phylum_glom, flist, TRUE)

speciesList <- tapply(sample_names(Age27_phylum_glom_MA), get_variable(Age27_phylum_glom_MA, "pH"), c)
speciesPhyseq <- lapply(speciesList, prune_samples, Age27_phylum_glom_MA)
speciesOTUtable <- lapply(speciesPhyseq,otu_table)
speciesAvg <- lapply(speciesOTUtable,rowMeans)
pooledOTUtable = t(do.call(rbind,speciesAvg))
pooledOTUtable = data.frame(OTU=row.names(pooledOTUtable),pooledOTUtable)

TT = tax_table(Age27_phylum_glom_MA)
TT = TT[, which(apply(!apply(TT, 2, is.na), 2, any))]
tdf = data.frame(TT, OTU = taxa_names(Age27_phylum_glom_MA))
pOTUtax = merge(pooledOTUtable, tdf, by.x = "OTU")
str(pOTUtax)
pOTU = data.frame(pOTUtax,SeqTotal = rowSums(pOTUtax[,2:4]))
pOTU = pOTU[order(-pOTU$SeqTotal),]

sum(pOTU$SeqTotal)/sum(pOTU$SeqTotal)

pOTU.phylum =pOTU[,c(2:4,6)]
melt.phylum = melt(pOTU.phylum,id.vars="Phylum")
colnames(melt.phylum)[2]="OTU"
agg.phylum=aggregate(.~Phylum+OTU,melt.phylum,sum)
revalue(agg.phylum$OTU, c("X7.4" = "7.4")) -> agg.phylum$OTU
revalue(agg.phylum$OTU, c("X7.8" = "7.8")) -> agg.phylum$OTU

d= ggplot(agg.phylum,aes(x=OTU,y=value,fill=Phylum)) +
  geom_bar(stat="identity",position="fill") +
  scale_y_continuous(labels = percent_format()) 
c= d+ xlab("pH")

b= c+ ylab("Relative Abundance")

a= b+ scale_fill_manual(name="Phylum",values = c("grey26", "royalblue", "chartreuse3",  "red", "darkorange","cyan2", "darkgreen", "deepskyblue", "mediumorchid3","#89C5DA", "#DA5724", "#74D944", "#C84248", "#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030",   "#AD6F3B", "#CD9BCD", "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#8569D5", "#5E738F", "#D1A33D", "#8A7C64", "#599861", "blue4", "yellow1", "violetred", "#990000", "#99CC00", "#003300", "#00CCCC", "#9966CC", "#993366", "#990033", "#4863A0", "#000033", "#330000", "#00CC99", "#00FF33", "#00CCFF", "#FF9933", "#660066", "#FF0066", "#330000", "#CCCCFF", "#3399FF", "#66FFFF", "#B5EAAA","#FFE87C"))

z= a+ theme(axis.title.x = element_text(face="bold",size=14),
            axis.text.x = element_text(angle=0, colour = "black", vjust=1, hjust = 1, size=14),
            axis.text.y = element_text(colour = "black", size=14),
            axis.title.y = element_text(face="bold",size=14),
            plot.title = element_text(size = 14),
            legend.title =element_text(face="bold",size = 14),
            legend.text = element_text(size = 14),
            legend.position="right",
            legend.key.size = unit(0.50, "cm"),
            strip.text.x = element_text(size=12, face="bold"),
            strip.text.y = element_text(size=12, face="bold"),
            panel.background = element_blank(),
            panel.border = element_rect(fill = NA, colour = "black"),
            strip.background = element_rect(colour="black"))

Fig2f= z+ guides(fill = guide_legend(reverse = TRUE, ncol=1))
Fig2f

ggsave("/file path/Github/Figure 2/Fig2f.pdf",width=6.5,height=4)
```

```{r Multiplot Taxa plot}
pdf("/file path/Github/Figure 2/Fig2_multitaxa.pdf",width=6,height=8)
multiplot(Fig2e,Fig2f, cols=1)
dev.off()

```
